---
title: "Getting Started with Recode Studio"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with Recode Studio}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

**Recode Studio** is a standalone, dataset-agnostic [Shiny](https://shiny.posit.co/) application and pure-R toolkit designed for cleaning messy string variables, grouping typo variants, building recode rules, and generating reproducible, deterministic data transformation pipelines.

This vignette demonstrates:
1. Identifying messy string columns and column metadata.
2. Grouping typo variants using string-similarity clustering.
3. Propagating recode rules across sibling columns (e.g., `drink1` and `drink2`).
4. Generating deterministic, single-pass `dplyr::case_when()` R code.

---

## 1. Inspecting Datasets and Free-Text Columns

Recode Studio includes a pure-R metadata scanner (`build_meta()`) that classifies columns into:
* **Identifiers/Numbers/Dates:** Automatically detected and excluded from text clustering.
* **Short text:** Categorical strings and labels suitable for similarity clustering and spellchecking.
* **Long text:** Paragraphs and multi-sentence reviews suitable for token analysis and word clouds.

```{r metadata}
# Load the bundled cafe dataset
csv_path <- system.file("extdata", "example_messy.csv", package = "recodeStudio")
if (!nzchar(csv_path)) csv_path <- file.path("..", "inst", "extdata", "example_messy.csv")

df <- readr::read_csv(csv_path, show_col_types = FALSE)
knitr::kable(head(df[, 1:6], 5))
```

We can inspect the column classifications:

```{r inspect_meta}
source(file.path("..", "R", "text_helpers.R"))
source(file.path("..", "R", "data_loader.R"))

meta <- build_meta(df)
knitr::kable(meta[, c("column", "is_free_text_candidate", "text_kind", "n_unique", "pct_na")])
```

---

## 2. String-Similarity Clustering

When inspecting short text columns like `drink1` (which contains variants such as `cappuccino`, `capuccino`, `cappucino`, `capccino`), Recode Studio groups them into similarity clusters.

```{r clustering}
source(file.path("..", "R", "string_helpers.R"))

# Cluster the unique values in drink1 using Jaro-Winkler distance
clusters <- cluster_strings(df$drink1, method = "jw", threshold = 0.85)

# View detected clusters
knitr::kable(head(clusters, 8))
```

---

## 3. Building Single-Pass Recode Rules

Recode rules are structured tibbles conforming to the canonical schema:
* `rule_id`: 12-character SHA-1 hash of `(variable, match_type, old_value)`.
* `variable`: Target column.
* `match_type`: `exact`, `exact_ci`, `trimmed_ci`, or `regex`.
* `old_value`: Pattern or string to replace.
* `new_value`: Standardized target value.
* `apply_to_siblings`: Boolean flag indicating whether to apply across sibling columns.
* `sibling_pattern`: Regex identifying sibling columns (e.g. `^drink[12]$`).

```{r rules}
rules <- tibble::tribble(
  ~rule_id, ~variable, ~apply_to_siblings, ~sibling_pattern, ~match_type, ~old_value, ~new_value, ~action, ~notes, ~author, ~created_at, ~updated_at, ~source_dataset,
  "rule_cap_01", "drink1", TRUE, "^drink[12]$", "trimmed_ci", "capuccino", "cappuccino", "recode", "typo fix", "reviewer", "2026-08-20", "2026-08-20", "example_messy",
  "rule_cap_02", "drink1", TRUE, "^drink[12]$", "trimmed_ci", "cappucino", "cappuccino", "recode", "typo fix", "reviewer", "2026-08-20", "2026-08-20", "example_messy",
  "rule_exp_01", "drink1", TRUE, "^drink[12]$", "trimmed_ci", "expresso",  "espresso",   "recode", "typo fix", "reviewer", "2026-08-20", "2026-08-20", "example_messy"
)

# Apply recodes directly to data
cleaned_df <- apply_recodes(df, rules)
knitr::kable(head(cleaned_df[, c("order_id", "drink1", "drink2")], 6))
```

---

## 4. Reproducible Code Generation

Recode Studio exports clean, standalone `dplyr` code that matches `apply_recodes()` with exact 1-to-1 agreement:

```{r codegen}
code <- generate_recode_R(rules, dataset_id = "df")
cat(code)
```

---

## Summary

Recode Studio allows teams to explore dirty datasets, build and validate rules interactively, and export self-contained, reproducible scripts for downstream reporting.
