---
title: "Long-Text Analysis & Interactive Visualizations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Long-Text Analysis & Interactive Visualizations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

Free-text variables like customer reviews, survey comments, and incident descriptions contain sentence- and paragraph-length content. For these columns, string-edit clustering is unhelpful.

Recode Studio's **Text analysis** module provides:
1. **Length distribution profiling** (character, word, and sentence counts).
2. **Top token and n-gram frequency extraction** with configurable stopword filtering.
3. **Keyword-In-Context (KWIC)** search.
4. **Interactive word clouds** via `wordcloud2`.

---

## 1. Profiling Long Text Columns

```{r load_data}
source(file.path("..", "R", "text_helpers.R"))

csv_path <- system.file("extdata", "example_survey.csv", package = "recodeStudio")
if (!nzchar(csv_path)) csv_path <- file.path("..", "inst", "extdata", "example_survey.csv")

df <- readr::read_csv(csv_path, show_col_types = FALSE)
comments <- df$detailed_comments

# Classify column length
classify_text_length(comments)

# Calculate summary metrics per response
stats <- text_length_stats(comments)
summary(stats)
```

---

## 2. Token & N-Gram Extraction

Extracting the most frequent individual words (unigrams) and phrases (bigrams/trigrams) while filtering common English stopwords:

```{r tokens}
# Top 10 words (excluding common English stopwords)
tokens_df <- top_tokens(comments, n = 10, remove_stopwords = TRUE, min_chars = 3)
knitr::kable(tokens_df)

# Top contiguous 2-word phrases
ngrams_df <- top_ngrams(comments, ngram = 2, n = 8, remove_stopwords = TRUE)
knitr::kable(ngrams_df)
```

---

## 3. Keyword in Context (KWIC)

KWIC lets reviewers inspect how specific words (e.g. `"performance"`, `"bug"`, `"support"`) are used across customer sentences:

```{r kwic}
kwic_results <- kwic(comments, term = "performance", window = 4)
knitr::kable(kwic_results)
```

---

## 4. Interactive Word Cloud Visualization

Recode Studio integrates `wordcloud2` to render HTML5 canvas token clouds directly in the Shiny UI:

```{r wordcloud, eval = FALSE}
# Inside Shiny text analysis server module:
df_cloud <- data.frame(
  word = as.character(tokens_df$token),
  freq = as.numeric(tokens_df$n),
  stringsAsFactors = FALSE
)

wordcloud2::wordcloud2(
  data = df_cloud,
  size = 0.6,
  color = "random-dark",
  backgroundColor = "white"
)
```

---

## Conclusion

Combining length distributions, n-gram mining, KWIC search, and interactive canvas visualizations allows rapid exploratory review of large unstructured text fields prior to recoding and categorization.
